---
title: "Predicting tree cover using landsate data"
author: "Mark  ten Vregelaar and Jos Goris"
date: "January 13, 2016"
output: html_document
---


## Set up the working environment
```{r, message=FALSE}
# Start with empty environment
rm(list=ls())

# Get required libraries
library(raster)
library(rgdal)
library(rasterVis)

# create data and output folder
ifolder <- "./data/"
ofolder <- "./output/"
dir.create(ifolder, showWarnings = FALSE)
dir.create(ofolder, showWarnings = FALSE)

```

## Downloading the required data and pre-processing
```{r,message=FALSE}
# Download and unzip landsate data-----------------------------------------------------------

dataURL <- "https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/archive/gh-pages.zip"
inputZip <- list.files(path=ifolder, pattern= '^.*\\.zip$')
if (length(inputZip) == 0){ ##only download when not alrady downloaded
	download.file(url = dataURL, destfile = 'data/landsate_data.zip', method = 'wget')
	
}
unzip('data/landsate_data.zip', exdir=ifolder)  # unzip NDVI data

# Load the data------------------------------------------------------------------------------
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB2.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB3.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB4.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/trainingPoly.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB1.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB5.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/GewataB7.rda")
load("data/AdvancedRasterAnalysis-gh-pages/data/vcfGewata.rda")

# Pre-processing------------------------------------------------------------------------------
# Assign NA to pixels with value >10
vcfGewata[vcfGewata>100]<- NA

# Organize the data into different variables 
vcf_B123 <- brick(vcfGewata,GewataB1,GewataB2,GewataB3)
vcf_B456 <- brick(vcfGewata,GewataB4,GewataB5,GewataB7)
allbands <- brick(vcfGewata,GewataB1,GewataB2,GewataB3,GewataB4,GewataB5,GewataB7)
VCF <- allbands[[1]]

bandnames <- c("VCF", "band1", "band2", "band3","band4", "band5", "band7")
names(allbands) <- bandnames
names(vcf_B123) <- bandnames[1:4]
names(vcf_B456) <- c(bandnames[1],bandnames[5:7])

# Make dataframe of allbands and omit NA
valuetable <- getValues(allbands)
valuetable <- na.omit(valuetable)
valuetable <- as.data.frame(valuetable)


```

## Create model for calculating VCF

Examining the correlations of the spectral bands with VCF
```{r,message=FALSE}

# Plot scatter plot of different spectral bands and VCF
pairs(vcf_B123)
pairs(vcf_B456)
```

Band 4 has very low correlation and is therefore not used in the model
```{r,message=FALSE}
# Calculate linear regression model for predicting VCF 
regmodel <- lm(VCF ~ band1 + band2 + band3 + band5 +band7, data=valuetable)
```

Summary of the model
```{r,message=FALSE}
# show model summary 
summary(regmodel)
```

Predict VCF using the linear model
```{r,message=FALSE}
# Use model to predict VCF values
predVCF <- predict(allbands, model=regmodel, na.rm=TRUE)

# Change values below 0 in predVCF to NA
predVCF[predVCF<0] <- NA
```



Evaluation of results by calculating difference and RMSE
```{r, message=FALSE}
# Calculate the differnce between predicted VCF and original VCF
difference <- (predVCF - VCF)
# Calculate RMSE
RMSE <- sqrt((cellStats((difference)^2,stat='mean', na.rm=TRUE)))
names(RMSE) <- 'RMSE total raster'
# Calculate RMSE for different zones
trainingPoly@data$Code <- as.numeric(trainingPoly@data$Class)
zones <- rasterize(trainingPoly, difference, field='Code')
diffmasked <- mask(difference, zones)
zoneRMSE <- sqrt(zonal(((diffmasked)^2), zones, fun='mean', digits=0, na.rm=TRUE))
rownames(zoneRMSE)<-c('cropland','forest','wetlands')
zoneRMSE<-zoneRMSE[,-1]

```

## Plot results
Predicted and priginal VCF
```{r,message=FALSE}
myColorkey <- list(at=seq(0, 100, 2.5), # where the colors change
                   labels=list(labels=seq(0, 100, 10),at=seq(0, 100, 10)# labels
                     ),space = "bottom")
levelplot(stack(VCF,predVCF), col.regions = colorRampPalette(c("brown",'yellow', "darkgreen"))(255),colorkey =myColorkey,names.attr=c('Original VCF','predicted VCF'),main=list('Tree cover in Gewata (Ethiopia)',cex=1.8,vjust=-1),scales=list(draw=FALSE),xlab=list('Tree cover (%)',vjust=7))

```

RMSE error

```{r, message=FALSE}
barplot(c(zoneRMSE,RMSE),main='RMSE per zone and total raster',ylab='RMSE',col=c("orange", "dark green", "light blue","blue"),ylim= c(0,12))
```

